---
title: "BS Hazard"
author: "Zachary Gager"
date: "11/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(ggplot2)
library(reshape2)
library(bshazard)
library(gridExtra)
source("../convert_list.R")
source("../SimulationFunctions.R")
```


# BSHazard for exponential hazard rate with sinusoidal fvt


Here, the $\textrm{mean}(\lambda_3(t)) = \textrm{mean}(\frac{0.05}{e} * e^{\textrm{fvt}_2(t)})$, which is just used to find the mean true hazard rate for all the currently alive individuals at each time step. 

### Simulate and Convert Data

The following simulates 1000 organisms in both cases where fvt influences or does not influence the hazard rate. The data also needed to be converted with ```convert_list.R``` for the ```bshazard``` function. Here we also found the average hazard rate at each time step.

```{r}
out = sim_exp_sin(1000)
fvt.data = out$fvt
nat.data = out$nat
# Calculate average hazard rate accross entire population 
# from each individual's fvt
true.fvt.hazard = as.vector(apply(fvt.data, 2, function(col) {
  col = na.omit(col)
  mean((0.05 / exp(1)) * exp(col))
}))[-1]

surv.data.fvt = convert_list(fvt.data)
surv.data.fvt$time0 = surv.data.fvt$time - 1

surv.data.nat = convert_list(nat.data)
surv.data.nat$time0 = surv.data.nat$time - 1
```

### Run ```bshazard```


```{r, results=F}
bs.fvt = bshazard(Surv(time0, time, status)~1, data=surv.data.fvt, phi=1)
bs.nat = bshazard(Surv(time0, time, status)~1, data=surv.data.nat, phi=1)
```

We also tested for ```Surv(time0, time, status)~trait```, but it takes much longer to run ```bshazard``` this way and there was not much of a difference in the results. 

### Make Plots

```{r}

nat.toplot = data.frame(
  "fit"=bs.nat$hazard, 
  "time"=bs.nat$time, 
  "true"=rep(0.05, length=length(bs.nat$time)),
  "lower.ci"=bs.nat$lower.ci,
  "upper.ci"=bs.nat$upper.ci
  )
fvt.toplot = data.frame(
  "fit"=bs.fvt$hazard, 
  "time"=bs.fvt$time, 
  "true"=true.fvt.hazard,
  "lower.ci"=bs.fvt$lower.ci,
  "upper.ci"=bs.fvt$upper.ci
  )

p1 = ggplot(nat.toplot) +
  geom_line(aes(x=time, y=fit, color="Fitted"), size=1) +
  geom_line(aes(x=time, y=true, color="Actual"), size=1) +
  geom_ribbon(aes(x=time, ymin=lower.ci, ymax=upper.ci), alpha=0.15) +
  theme_bw() +
  scale_color_manual("", breaks=c("Fitted", "Actual"), values=c("Fitted"="black", "Actual"="red")) +
  ggtitle("Trait Does Not Influence Survival") +
  xlab("Time") +
  ylab("Hazard Rate") +
  theme(plot.title = element_text(size=18, hjust=0.5), axis.title=element_text(size=12), legend.text=element_text(size=10))

p2 = ggplot(fvt.toplot) +
  geom_line(aes(x=time, y=fit, color="Fitted"), size=1) +
  geom_line(aes(x=time, y=true, color="Actual"), size=1) +
  geom_ribbon(aes(x=time, ymin=lower.ci, ymax=upper.ci), alpha=0.15) +
  theme_bw() +
  scale_color_manual("", breaks=c("Fitted", "Actual"), values=c("Fitted"="black", "Actual"="red")) +
  ggtitle("Trait Does Influence Survival") +
  xlab("Time") +
  ylab("Hazard Rate") +
  theme(plot.title = element_text(size=18, hjust=0.5), axis.title=element_text(size=12), legend.text=element_text(size=10))

grid.arrange(p1,p2, nrow=2)

```


Above, the actual hazard rate become more varied towards the end because there are less organisms alive, so the average hazard rate has a greater bias towards specific organisms.


# BSHazard for linear hazard rate with Holling type 2

Here, the $\textrm{mean}(\lambda_1(t)) = \textrm{mean}(0.05 + 0.001 * \textrm{fvt}_1(t))$.


### Simulate and Convert Data

Like above, except we simulate for Holling Type II, and calculate the average hazard rate of the population differently. 

```{r}
out = sim_hollings(1000)
fvt.data = out$fvt
nat.data = out$nat
# Calculate average hazard rate accross entire population 
# from each individual's fvt
true.fvt.hazard = as.vector(apply(fvt.data, 2, function(col) {
  col = na.omit(col)
  mean(0.05 + 0.001* col)
}))[-1]

surv.data.fvt = convert_list(fvt.data)
surv.data.fvt$time0 = surv.data.fvt$time - 1

surv.data.nat = convert_list(nat.data)
surv.data.nat$time0 = surv.data.nat$time - 1

```

### Run ```bshazard```

```{r, results=F}
bs.fvt = bshazard(Surv(time0, time, status)~1, data=surv.data.fvt, phi=1)
bs.nat = bshazard(Surv(time0, time, status)~1, data=surv.data.nat, phi=1)
```


### Make Plots

```{r}
nat.toplot = data.frame(
  "fit"=bs.nat$hazard, 
  "time"=bs.nat$time, 
  "true"=rep(0.05, length=length(bs.nat$time)),
  "lower.ci"=bs.nat$lower.ci,
  "upper.ci"=bs.nat$upper.ci
  )
fvt.toplot = data.frame(
  "fit"=bs.fvt$hazard, 
  "time"=bs.fvt$time, 
  "true"=true.fvt.hazard,
  "lower.ci"=bs.fvt$lower.ci,
  "upper.ci"=bs.fvt$upper.ci
  )

p1 = ggplot(nat.toplot) +
  geom_line(aes(x=time, y=fit, color="Fitted"), size=1) +
  geom_line(aes(x=time, y=true, color="Actual"), size=1) +
  geom_ribbon(aes(x=time, ymin=lower.ci, ymax=upper.ci), alpha=0.15) +
  theme_bw() +
  scale_color_manual("", breaks=c("Fitted", "Actual"), values=c("Fitted"="black", "Actual"="red")) +
  ggtitle("Trait Does Not Influence Survival") +
  xlab("Time") +
  ylab("Hazard Rate") +
  theme(plot.title = element_text(size=18, hjust=0.5), axis.title=element_text(size=12), legend.text=element_text(size=10))

p2 = ggplot(fvt.toplot) +
  geom_line(aes(x=time, y=fit, color="Fitted"), size=1) +
  geom_line(aes(x=time, y=true, color="Actual"), size=1) +
  geom_ribbon(aes(x=time, ymin=lower.ci, ymax=upper.ci), alpha=0.15) +
  theme_bw() +
  scale_color_manual("", breaks=c("Fitted", "Actual"), values=c("Fitted"="black", "Actual"="red")) +
  ggtitle("Trait Does Influence Survival") +
  xlab("Time") +
  ylab("Hazard Rate") +
  theme(plot.title = element_text(size=18, hjust=0.5), axis.title=element_text(size=12), legend.text=element_text(size=10))
grid.arrange(p1,p2)
```








