---
title: "Spline_Model_Results"
author: "Ben Browder"
date: "12/13/2021"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(ModelMetrics)
source("../convert_list.R")
source("x_transform_data.R")
source("spline_model_functions.R")
source("../SimulationFunctions.R")
source("../f_beta_eval.R")
```

## Creating Control and Experimental Holling Linear Interpolated Datasets in order to train on.
```{r}
holling_data <- x_transform_data(read.csv('../Data/fvt_1000.csv'))
holling_data1 <- x_transform_data(read.csv('../Data/natural_1000.csv'))
```

## Train function to get Spline average intercept, coefficients and threshold value from Holling Control and Experimental datasets combined.
```{r}
results <- spline_coefficient_train(holling_data, holling_data1)
results
```

## Testing Binary results using function with parameters.
```{r}
spline_coefficient_test_params(read.csv('../Data/natural_1000.csv'), results$int, results$df1, results$df2, results$df3, results$df4, results$thresh)
```
## Testing Binary results using function with no parameters.
```{r}
spline_coefficient_test(read.csv('../Data/natural_1000.csv'))
```

## Testing Binary results using Accuracy test of 30 organisms with a total of 100 Holling datasets, 50 control and 50 experimental.
```{r}
coef_test <- f_beta_eval(spline_coefficient_test, sim_hollings, 30, 50)
coef_test
```

## Testing Binary results using Accuracy test of 50 organisms with a total of 100 Holling datasets, 50 control and 50 experimental.
```{r}
coef_test <- f_beta_eval(spline_coefficient_test, sim_hollings, 50, 50)
coef_test
```

## Testing Binary results using Accuracy test of 100 organisms with a total of 100 Holling datasets, 50 control and 50 experimental.
```{r}
coef_test <- f_beta_eval(spline_coefficient_test, sim_hollings, 100, 50)
coef_test
```

## Creating Control and Experimental Sinusodal Linear Interpolated Datasets in order to train on.
```{r}
sin_data <- x_transform_data(read.csv('../Data/sin_linearhazard_control_1000org.csv'))
sin_data1 <- x_transform_data(read.csv('../Data/sin_linearhazard_experimental_1000org.csv'))
```

## Train function to get Spline average intercept, coefficients and threshold value from Sinusodal Control and Experimental datasets combined.
```{r}
results <- spline_coefficient_train(sin_data, sin_data1)
results
```
## Testing Binary results using Accuracy test of 30 organisms with a total of 100 Sinusodal datasets, 50 control and 50 experimental.
```{r}
coef_test <- f_beta_eval(spline_coefficient_test, sim_sin, 30, 50)
coef_test
```

## Rest of results for Sinusodal return 50% accuracy with F-Beta score of 0.
