---
title: "Original Spline plots"
author: "Ben Browder"
date: "12/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(ggplot2)
library(dplyr)
library(ggfortify)
source("convert_list.R")
source("Splines/x_transform_data.R")
source("Splines/x_transform_plot.R")
source("Splines/spline_fits.R")
```

## Creating Linear Interpolated datasets of control and experimental group
```{r}
Trait_Influences_Survival <- x_transform_data(read.csv('Data/fvt.csv'))
Trait_Does_Not_Influence_Survival <- x_transform_data(read.csv('Data/natural.csv'))
```

## Plot of Linear Interpolated experimental group
```{r}
x_transform_plot(Trait_Influences_Survival)
```

## Plot of Linear Interpolated control group
```{r}
x_transform_plot(Trait_Does_Not_Influence_Survival)
```
## Plot of linearly interpolated experimental dataset with average Spline fit as well as a linear regression table of the Spline fit.
```{r}
spline_fits(Trait_Influences_Survival)
```
## Plot of linearly interpolated control dataset with average Spline fit as well as a linear regression table of the Spline fit.
```{r}
spline_fits(Trait_Does_Not_Influence_Survival)
```